### 2018-06-16 Paper/examples.tex
..

%
Cheap function evaluations of the discharge $\mathcal{Q}(\vec{\theta})$ 
enables computing accurate estimates of $\mathcal{T}(\theta_i)$ through
sampling, with a sufficiently large MC sample size.  Shown in
Figure~\ref{fig:sense_bore}~(left) are estimates of these indices corresponding to the
uncertain parameters in the borehole function using 10$^6$ pseudo-random
samples in the input parameter space. 
%\footnote{Although $\mathcal{T}(\theta_i)$ might converge with much fewer
%samples depending upon the model, we consider a large number that typically
%ensures a converged estimate for the purpose of illustration.} 
These estimates are used to verify fidelity of
parameter screening based on the methodology presented in Section~\ref{sec:method}.

@@

As discussed earlier in this section, the reduced-space PCE's are verified for
predictive accuracy in a least-squares sense and a probabilistic sense.
Estimates for $\epsilon_{\mbox{\tiny{L-2}}}$ based on 50 samples in the
validation test suite were
found to be 0.0551 and 0.0112 for the 4D and 5D PCE's, respectively. In other
words, the 4D PCE is accurate within 5.52$\%$ and the 5D PCE is accurate within
1.12$\%$ of predictions based on the borehole function. We note  that although
the convergence in the case of a 5D PCE is slower (Figure~\ref{fig:conv_bore}),
its predictive accuracy is higher than the 4D PCE. This illustrates the
trade-off between accuracy and computational efficiency for the present 
problem. Generally, the required level of accuracy is problem dependent. 
The present framework allows for moving towards higher fidelity 
reduced-space surrogates based on the ranking of the parameter sensitivities.

@@

We aim to construct a reduced-space surrogate for the following QoI:
\be
\mathcal{F}(\vec\theta) = \frac{1}{|D|} \int_D u(\vec{x}; \vec\theta) \, d\vec{x}, 
\label{eq:qoi}
\ee
%
where $D$ is the region $[2/5, 3/5] \times [2/5, 3/5] \subset \Omega$, 
and $|D|$ denotes the area of $D$. 
While this model is considerably more complex than the previous numerical
examples, it can still be solved efficiently.
The equation was discretized using finite differences, and Newton's method
was used to solve the resulting nonlinear system on a $100 \times 100$ 2D
cartesian grid.
We computed converged estimates of the Sobol
total-effect index $\mathcal{T}(\theta_i)$, reported 
in Figure~\ref{fig:sense_elliptic}~(left) by solving the above non-linear
system at 500 samples in the 10-dimensional parameter space. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.46\textwidth]{./Figures/sense_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_elliptic}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the 
semilinear elliptic PDE~\eqref{eq:semilinear}. Right: 
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) for each uncertain parameter,
obtained using $N$ = 5, 10, 15, and 20 samples in the full parameter space.}
\label{fig:sense_elliptic}
\end{center}
\end{figure}
%
Sensitivity predictions based on the screening metric, $\widehat{\mathcal{C}_i\mu_i}$,
plotted in Figure~\ref{fig:sense_elliptic}~(right), are found to
be in close agreement with $\mathcal{T}(\theta_i)$, even for the case when $N$ = 5. As $N$
is increased from 5 to 20, estimates of the screening metric are observed to converge.
Based on the trends observed in Figure~\ref{fig:sense_elliptic}, it can be said that
the uncertainty in the QoI in \eqref{eq:qoi} is largely dependent on $c$, 
$\alpha_2$, $\alpha_3$, $\alpha_4$, and $\alpha_5$. These observations underscore the
potential for computational gains by constructing an RSS in the 5D parameter space. We 
illustrate the comparison of convergence characteristics of the PCEs constructed in the
full parameter space (10D) and the reduced space (5D) in Figure~\ref{fig:conv_elliptic} (left). 
As expected, the RSS converges considerably faster. Using model evaluations at 90
sample points, $\epsilon_{\mbox{\tiny{LOO}}}$ is found to be two orders of magnitude
smaller than that in the case of full-surrogate ($\mathcal{O}(10^{-4}$) versus
$\mathcal{O}(10^{-2}$)). 
Consequently, the computational effort for constructing the RSS in the present test problem
is expected to be much smaller. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/err_samples_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_elliptic}
\caption{Left: Logarithm of $\epsilon_{\mbox{\tiny{LOO}}}$ is plotted against sample size for 
PCEs constructed in 10 and 5 dimensions to compare their convergence characteristics. 
Right: PDF of the QoI, $\mathcal{F}(\vec\theta)$ in~\eqref{eq:semilinear} is
plotted using the full-surrogate and the reduced-space PC surrogate.}
\label{fig:conv_elliptic}
\end{center}
\end{figure}

>>>>

%
Cheap function evaluations of the discharge $\mathcal{Q}(\vec{\theta})$ 
enables construction of the FSS with minimal effort. FSS predictions at
a large set of MC samples in the input space are used to obtain converged
estimates of $\mathcal{T}(\theta_i)$. Shown in
Figure~\ref{fig:sense_bore}~(left) are estimates of these indices corresponding to the
uncertain parameters in the borehole function using 10$^6$ MC
samples in the input parameter space. 
%\footnote{Although $\mathcal{T}(\theta_i)$ might converge with much fewer
%samples depending upon the model, we consider a large number that typically
%ensures a converged estimate for the purpose of illustration.} 
These estimates are used to verify fidelity of
parameter screening based on the methodology presented in Section~\ref{sec:method}.

@@

As discussed earlier in this section, the reduced-space PCE's are verified for
predictive accuracy in a least-squares sense and a probabilistic sense.
Estimates for $\epsilon_{\mbox{\tiny{L-2}}}$ based on 50 samples in the
validation test suite were
found to be 0.0551 and 0.0112 for the 4D and 5D PCE's, respectively. In other
words, the 4D PCE is accurate within 5.52$\%$ and the 5D PCE is accurate within
1.12$\%$ of predictions based on the borehole function. Note that $\epsilon_\text{\tiny{LOO}}$
however, is lower in the case of 4D PCE (Figure~\ref{fig:conv_bore}).
This illustrates the
trade-off between accuracy and computational efficiency for the present 
problem. Generally, the required level of accuracy is problem dependent. 
The present framework allows for moving towards higher fidelity 
reduced-space surrogates based on the ranking of the parameter sensitivities.

@@

We aim to construct a reduced-space surrogate for the following QoI:
\be
\mathcal{F}(\vec\theta) = \frac{1}{|D|} \int_D u(\vec{x}; \vec\theta) \, d\vec{x}, 
\label{eq:qoi}
\ee
%
where $D$ is the region $[2/5, 3/5] \times [2/5, 3/5] \subset \Omega$, 
and $|D|$ denotes the area of $D$. 
While this model is considerably more complex than the previous numerical
examples, it can still be solved efficiently.
The equation was discretized using finite differences, and Newton's method
was used to solve the resulting nonlinear system on a $100 \times 100$ 2D
cartesian grid.
We computed converged estimates of the Sobol
total-effect index $\mathcal{T}(\theta_i)$, reported 
in Figure~\ref{fig:sense_elliptic}~(left) using FSS predictions at 10$^6$
MC samples in the input space. The FSS was 
constructed using model predictions at 500 training points in the 10-dimensional
input space. Corresponding value of $\epsilon_\text{\tiny{LOO}}$ was 
found to be 9.729$\times 10^{-4}$. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.46\textwidth]{./Figures/sense_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/ub_conv_elliptic}
\caption{
Left: Sobol' total sensitivity index, $\mathcal{T}(\theta_i)$ for uncertain parameters in the 
semilinear elliptic PDE~\eqref{eq:semilinear}. Right: 
Estimates of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$) for each uncertain parameter,
obtained using $N$ = 5, 10, 15, and 20 samples in the full parameter space.}
\label{fig:sense_elliptic}
\end{center}
\end{figure}
%
Sensitivity predictions based on the screening metric, $\widehat{\mathcal{C}_i\mu_i}$,
plotted in Figure~\ref{fig:sense_elliptic}~(right), are found to
be in close agreement with $\mathcal{T}(\theta_i)$, even for the case when $N$ = 5. As $N$
is increased from 5 to 20, estimates of the screening metric are observed to converge.
Based on the trends observed in Figure~\ref{fig:sense_elliptic}, it can be said that
the uncertainty in the QoI in \eqref{eq:qoi} is largely dependent on $c$, 
$\alpha_2$, $\alpha_3$, $\alpha_4$, and $\alpha_5$. These observations underscore the
potential for computational gains by constructing an RSS in the 5D parameter space. We 
illustrate the comparison of convergence characteristics of the PCEs constructed in the
full parameter space (10D) and the reduced space (5D) in Figure~\ref{fig:conv_elliptic} (left). 
As expected, the RSS converges considerably faster. Using model evaluations at 90
sample points, $\epsilon_{\mbox{\tiny{LOO}}}$ is found to be two orders of magnitude
smaller than that in the case of full-surrogate ($\mathcal{O}(10^{-4}$) versus
$\mathcal{O}(10^{-2}$)). 
Consequently, the computational effort for constructing the RSS in the present test problem
is expected to be much smaller. 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/err_samples_elliptic}
  \includegraphics[width=0.48\textwidth]{./Figures/pdf_comp_elliptic}
\caption{Left: Logarithm of $\epsilon_{\mbox{\tiny{LOO}}}$ is plotted against sample size for 
PCEs constructed in 10 and 5 dimensions to compare their convergence characteristics. 
Right: PDF of the QoI, $\mathcal{F}(\vec\theta)$ in~\eqref{eq:semilinear} is
plotted using the full-surrogate and the reduced-space PC surrogate.}
\label{fig:conv_elliptic}
\end{center}
\end{figure}
